Viscous Erosion in Porous Media

### 2018-01-10 docs/methods/erosion_methods.tex
NM

% BRYAN'S ORIGINAL TEXT
%In this section, we also describe our time stepping method and two regularization terms that are necessary to control the curvature of the bodies and to maintain spectral accuracy in space.

%Once the shear stress is computed, it is used to update the shape of the bodies according to the normal flow~\eqref{eqn:bodyEvolution}.  To eliminate tangling and stretching of the mesh, we use a {\thL}formulation~\cite{hou-low-she1994} which is described in Section~\ref{sec:thetaL}.

%We use a quasi-static approximation so that the velocity field instantaneously changes with the changing geometry.  Therefore, we can alternate between the two main aspects of our simulation
%\begin{enumerate}
%\item Given a geometry, compute the shear stress of along the boundary of each of the bodies.
%\item Given the shear stress, update the geometry according to the erosion model.
%\end{enumerate}

@@

@@

First, we introduce the variables
\begin{align}
\label{eq:M}
& \Mterm = - 2 \pi \mean{\thalpha \Vn} \\
\label{eq:N}
& \NLterm = \frac{2 \pi}{L} \left( \pderiv{\atau}{\alpha} + \thalpha \Vs \right) \\
\label{eq:elfun}
& \elfun = \frac{2 \pi  \mean{\atau}}{L}
\end{align} 
so that evolution equations \eqref{eqn:Levo1} and \eqref{eqn:thetaevo1} can be compactly expressed as
\begin{align}
\label{eqn:Levo2}
& \tderiv{L}{t} = \Mterm \\
\label{eqn:thetaevo2}
& \pderiv{\theta}{t} = \eps \elfun \ppd{\theta}{\alpha} + \NLterm
\end{align}
We will evolve $\theta$ in spectral space, and so introduce the transforms
\begin{align}
& \theta(\alpha,t) = \alpha + \FourierSum \thhat_k e^{ i k \alpha} \\
& \NLterm(\alpha,t)  = \FourierSum \hat{\NLterm}_k e^{ i k \alpha}
\end{align}
Recall that $\alpha = 2 \pi s / L$ is a normalized arc length. In spectral space, \eqref{eqn:thetaevo2} produces the system of ODEs
\begin{equation}
\tderiv{\thhat_k}{t} +  \eps k^2  \elfun \thhat_k = \hat{\NLterm}_k
\end{equation}
Each ODE is first-order and linear (though variable-coefficient due to $\elfun$) and thus can be solved by the integrating factor
\begin{equation}
\label{eq:mu}
\mu_k = \exp \left( \eps k^2 \int \elfun \, dt \right)
\end{equation}
We are therefore led to the coupled system of ODEs
\begin{align}
\label{eqn:Levo2}
& \tderiv{L}{t} = \Mterm \\
\label{eqn:thetaevo2}
& \tderiv{}{t}\left( \mu_k \thhat_k \right) = \mu_k \hat{\NLterm}_k
\end{align}
We solve system \eqref{eqn:Levo2}--\eqref{eqn:thetaevo2} with an explicit, second-order Runge-Kutta method (RK2). In particular, we choose the {\em midpoint} RK2 method so as to maximize the smoothing effect of the curvature term.

>>>>

% Quasi-steady comment
%We use a quasi-static approximation so that the velocity field instantaneously changes with the changing geometry.  Therefore, we can alternate between the two main aspects of our simulation
%\begin{enumerate}
%\item Given a geometry, compute the shear stress of along the boundary of each of the bodies.
%\item Given the shear stress, update the geometry according to the erosion model.
%\end{enumerate}

@@

For the sake of numerical stability, we will regularize the boundary evolution in two ways. First, we will smooth the absolute

@@

First, we introduce the variables
\begin{align}
\label{eq:M}
& \Mterm = - 2 \pi \mean{\thalpha \Vn} \\
\label{eq:N}
& \NLterm = \frac{2 \pi}{L} \left( \pderiv{\atau}{\alpha} + \thalpha \Vs \right) \\
\label{eq:elfun}
& \elfun = \frac{2 \pi  \mean{\atau}}{L}
\end{align} 
so that evolution equations \eqref{eqn:Levo1} and \eqref{eqn:thetaevo1} can be compactly expressed as
\begin{align}
\label{eqn:Levo2}
& \tderiv{L}{t} = \Mterm \\
\label{eqn:thetaevo2}
& \pderiv{\theta}{t} = \eps \elfun \ppd{\theta}{\alpha} + \NLterm
\end{align}
Here $\NLterm$ represents the nonlinear terms (and nonlocal etc) in the PDE for $\theta$.
We will evolve $\theta$ in spectral space, and so introduce the transforms
\begin{align}
& \theta(\alpha,t) = \alpha + \FourierSum \thhat_k e^{ i k \alpha} \\
& \NLterm(\alpha,t)  = \FourierSum \hat{\NLterm}_k e^{ i k \alpha}
\end{align}
Recall that $\alpha = 2 \pi s / L$ is a normalized arc length. In spectral space, \eqref{eqn:thetaevo2} produces the system of ODEs
\begin{equation}
\tderiv{\thhat_k}{t} +  \eps k^2  \elfun \thhat_k = \hat{\NLterm}_k
\end{equation}
Each ODE is first-order and linear (though variable-coefficient due to $\elfun$) and thus can be solved by using the integrating factor
\begin{equation}
\label{eq:mu}
\mu_k = \exp \left( \eps k^2 \int \elfun \, dt \right)
\end{equation}
We are therefore led to the coupled system of ODEs
\begin{align}
\label{eqn:Levo2}
& \tderiv{L}{t} = \Mterm \\
\label{eqn:thetaevo2}
& \tderiv{}{t}\left( \mu_k \thhat_k \right) = \mu_k \hat{\NLterm}_k
\end{align}
We solve system \eqref{eqn:Levo2}--\eqref{eqn:thetaevo2} with an explicit, second-order Runge-Kutta method (RK2). In particular, we choose the {\em midpoint} RK2 method so as to maximize the smoothing effect of the curvature term.
